WDIAS : A Microservices-Based Weather Data Integration and Assimilation System\\

### 2020-05-08 method/sec_4-db_struct.tex
v5.1.1

Based on the above interpretation, a timeseries have data points ordered in time. Other than that, timeseries have additional information that is known as \emph{timeseries metadata}, which can use to identify one timeseries from another uniquely. As an example, we can place a weather station on a \emph{location}, and it can measure different types of weather data that are called \emph{parameters} such as temperature, precipitation, and wind direction. Each parameter can produce different timeseries. Even we can use the location as metadata to identify the timeseries, but that is not enough to distinctly identify the parameters. Thus, we need to use a set of attributes to identify a timeseries uniquely. In the next \cref{subse:timeseries_key_attributes}, we describe key attributes that are used by the \acrshort{wdias} to identify a timeseries uniquely.

@@

\paragraph{Value Type}-- Scalar, Vector, Grid are the Value Types that are interested in the \acrshort{wdias}.

@@

\begin{itemize}  
  \item \emph{Variable} -- Nature of the variable measuring. Example Precipitation, Temperature and Water Level etc
  \item \emph{Unit} -- metric unit of measuring
  \item \emph{Parameter Type} -- Should be one of Instantaneous, Accumulative, or Mean
\begin{lstlisting}[language=Python]
{
  "parameterId": "O.Precipitation",
  "variable": "Precipitation",
  "unit": "mm",
  "parameterType": "Instantaneous"
}
\end{lstlisting}
\end{itemize}

@@

We used SQL as the \acrshort{rdbms}. Since one of \acrshort{wdias} main goals is to use open source tools, it used MySQL as a persistent database for storing timeseries metadata described above. One MySQL instance used only for adapter-metadata service. Other than that, another MySQL instance used for adapter-extension. We will explain in more detail in \cref{se:data_preprocess}. We used MySQL only for the above services, and those database instances keep the consistency of the schema for timeseries metadata and extension metadata. Further, we used an In-Memory database instance for caching the requested metadata to reduce the load on the MySQL database instances while repeated queries. Other than MySQL, many open-source SQL databases can be used to replace the same functionality.

@@

We used Redis \cite{redisRedisDocumentation} as the In-Memory database since it is a widely used database system with excellent support. It also supports Publisher Subscriber capabilities as an inbuilt feature. Redis is an open-source (BSD licensed). Clustering is available for scalability. Redis uses in-memory caching for fast access to the frequent access data in adapter-metadata and adapter-extension. Other than that, adapter-status uses Redis as a key-value pair store to maintain the status for async requests handling while storing grid data. The adapter-extension uses Redis's Publisher Subscriber capabilities to create a cronjob in extension-scheduler when a new extension creates with a trigger on time during data preprocessing.

>>>>

Based on the above interpretation, a timeseries have data points ordered in time. Other than that, timeseries have additional information that is known as \emph{timeseries metadata}, which can use to identify one timeseries from another uniquely. As an example, we can place a weather station on a \emph{location}, and it can measure different types of weather data that are called \emph{parameters} such as temperature, precipitation, and wind direction. Each parameter can produce different timeseries. Even we can use the location as metadata to identify the timeseries, but that is not enough to distinctly identify the parameters. Thus, we need to use a set of attributes to identify a timeseries uniquely. Next, we describe key attributes that are used by the \acrshort{wdias} to identify a timeseries uniquely.

@@

\paragraph{Value Type}-- Scalar, vector, grid are the Value Types that are interested in the \acrshort{wdias}.

@@

\begin{itemize}  
  \item \emph{Variable} -- Nature of the variable measuring. Example precipitation, temperature and water level
  \item \emph{Unit} -- metric unit of measuring
  \item \emph{Parameter Type} -- Should be one of Instantaneous, Accumulative, or Mean
\begin{lstlisting}[language=Python]
{
  "parameterId": "O.Precipitation",
  "variable": "Precipitation",
  "unit": "mm",
  "parameterType": "Instantaneous"
}
\end{lstlisting}
\end{itemize}

@@

Since one of \acrshort{wdias} main goals is to use open source tools, it used MySQL as a persistent database for storing timeseries metadata described above. One MySQL instance used only for adapter-metadata service. Other than that, another MySQL instance used for adapter-extension. We will explain in more detail in \cref{se:data_preprocess}. We used MySQL only for the above services, and those database instances keep the consistency of the schema for timeseries metadata and extension metadata. Further, we used an In-Memory database instance for caching the requested metadata to reduce the load on the MySQL database instances while repeated queries. Other than MySQL, many open-source SQL databases can be used to replace the same functionality.

@@

We used Redis \cite{redisRedisDocumentation} as the In-Memory database since it is a widely used database system with excellent support. It also supports Publisher Subscriber capabilities as an inbuilt feature. Redis is an open-source. Clustering is available for scalability. Redis uses in-memory caching for fast access to the frequent access data in adapter-metadata and adapter-extension. Other than that, adapter-status uses Redis as a key-value pair store to maintain the status for async requests handling while storing grid data. The adapter-extension uses Redis's Publisher Subscriber capabilities to create a cronjob in extension-scheduler when a new extension creates with a trigger on time during data preprocessing.

### 2020-05-08 method/sec_5-data_preprocess.tex
v5.1.1

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.5\textwidth]{method/data_preprocess/weather_data_preprocessing.jpg}
    \caption{A generic mathematical function model for weather data preprocessing}
    \label{fi:weather_data_preprocessing}
\end{figure}

The data preprocess functionalities support simple generic mathematical function models. Each extension is considered as a mathematical function that can take \texttt{p} number of input timeseries variables and output \texttt{n} number of timeseries variables after applying the given function. Further, the system supports to provide a bind constant as configurations to the extension at the time of preprocessing. Thus, users can change the behavior of an existing function by providing different bind constants at the time of creating new triggers for an extension and produce different behaviors.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{method/data_preprocess/serial_interpolation.jpg}
    \caption{Serial interpolation.}
    \label{fi:serial_interpolation}
\end{figure}

\paragraph{Spatial Interpolation}-- During spatial interpolation, we use interpolation to either fill gaps in timeseries or to create a new timeseries for a location using data from other (spatially distributed) locations. Spatial interpolation can also be applied for sampling scalar timeseries from grid timeseries, for re-sampling grids, or for creating grids from timeseries data.
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{method/data_preprocess/spatial_interpolation.jpg}
    \caption{Spatial interpolation.}
    \label{fi:spatial_interpolation}
\end{figure}

\subsection{Transformation}
Simple arithmetic manipulation, Time interval transformation and Shifting the series in time Specific hydro-meteorological transformation such as stage discharge relationships.
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{method/data_preprocess/transformation.jpg}
    \caption{Time interval aggregration.}
    \label{fi:transformation}
\end{figure}

\subsection{Validation}
Checks for counting reliable, doubtful, unreliable, and missing values
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{method/data_preprocess/validation.jpg}
    \caption{Timeseries data validation.}
    \label{fi:validation}
\end{figure}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{method/data_preprocess/validation.jpg}
    \caption{Extension handler.}
    \label{fi:extension_handler}
\end{figure}

>>>>

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{method/data_preprocess/weather_data_preprocessing.pdf}
    \caption{A generic mathematical function model for weather data preprocessing}
    \label{fi:weather_data_preprocessing}
\end{figure}

The data preprocess functionalities support simple generic mathematical function models. Each extension is considered as a mathematical function that can take \texttt{p} input timeseries variables and output \texttt{n} timeseries variables after applying the given function. Further, the system supports to provide a bind constant as configurations to the extension at the time of preprocessing. Thus, users can change the behavior of an existing function by providing different bind constants at the time of creating new triggers for an extension and produce different behaviors.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{method/data_preprocess/serial_interpolation.pdf}
    \caption{Serial interpolation.}
    \label{fi:serial_interpolation}
\end{figure}

\paragraph{Spatial Interpolation}-- During spatial interpolation, we use interpolation to either fill gaps in timeseries or to create a new timeseries for a location using data from other (spatially distributed) locations. Spatial interpolation can also be applied for sampling scalar timeseries from grid timeseries, for re-sampling grids, or for creating grids from timeseries data.
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{method/data_preprocess/spatial_interpolation.pdf}
    \caption{Spatial interpolation.}
    \label{fi:spatial_interpolation}
\end{figure}

\subsection{Transformation}
Simple arithmetic manipulation, Time interval transformation and Shifting the series in time Specific hydro-meteorological transformation such as stage discharge relationships.
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{method/data_preprocess/transformation.pdf}
    \caption{Time interval aggregration.}
    \label{fi:transformation}
\end{figure}

\subsection{Validation}
Checks for counting reliable, doubtful, unreliable, and missing values
\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{method/data_preprocess/validation.pdf}
    \caption{Timeseries data validation.}
    \label{fi:validation}
\end{figure}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{method/data_preprocess/validation.pdf}
    \caption{Extension handler.}
    \label{fi:extension_handler}
\end{figure}

### 2020-05-08 results/sec_1-test_plan.tex
Added fixes
Error Term,Correct Term,Description,No of Occurrences,Pages (absolute)
a.k.a.,aka.,,Shoudl be aka and need a comma,1,-1
[a-z] as well as,, as well as,Needs a comma before,3,-1
Also,Also,,Comma should be used at beginning of a sentence,1,-1
can not,cannot,Is a single word,4,-1
[fF]igure.\d,Figure n,There has to be a space like Figure 2.7,82,-1
Hence,Hence,,Comma should be used at beginning of a sentence,1,-1
in order to,to,Can be replaced with just to in many cases,4,-1
in to,into,Is a single word,1,-1
may be,maybe,Is a single word,2,-1
section,Section,S should be capital if used li

\section{Load Testing Plan}
\label{se:test_plan}

In \cref{subse:modules_weather_data_integration_sys} we discussed three modules of weather data integration and assimilation systems such as integration, assimilation, and dissemination. We can map each of those modules to WDIAS modules, which provide the same functionalities as below.
\begin{enumerate}
    \item \emph{Import} modules - Integration modules
    \item \emph{Export} modules - Dissemination modules
    \item \emph{Extension} modules - Assimilation modules
\end{enumerate}

@@

A timeseries is a series of data points in time order. The system may receive the data points of a timeseries in multiple requests. The request size can vary based on the use case from one data point to multiple data points. For example;
\begin{itemize}
    \item A weather station may send the current water level in a regular period of 1 minute. In such a case, it always comes as a request with one data point with a higher frequency.
    \item Due to resource issues such as batteries, a weather station can collect data points for a given period and send as a batch. An example scenario would be, a weather station can collect the precipitation at each 5 minute period and send all the data after collecting for one hour.
    \item The forecast precipitation data from the \acrshort{wrf} model extract for a few days in hourly intervals and store them.
\end{itemize}

The percentage of number of requests per each data type that comes to the system at a given time can be varied depending on the situation. The Majority of weather data requests come from weather stations and other observation tools. Then we need to validate and modify those timesereis data and store them as new timeseries that can directly feed into the NWMs.. These weather stations can send the weather data at different frequencies with different parameters such as precipitation, temperature, humidity, and wind direction. Most of the timeseries mentioned above are scalar timeseries. When compared to scalar data, parameters such as wind direction are vector timeseries with a lesser percentage of requests. On the other hand, most of the grid type data was generated as results of NWMs. But those grid data produce in less frequency due to the cost of the computer resource to produce results. After considering those factors, we came up with 70\% scalar, 20\% vector, and 10\% grid percentage values at the expected requests rate ratios.
As an upper limit of such scenarios while designing the performance testing, it uses the following;
\begin{itemize}
    \item Hourly (24 data points)
    \item Every 30 minutes (48 data points)
    \item Every 15 minutes (96 data points) - One request is equivalent to a set of data points sent by a weather station in 15 minutes interval for a single day
\end{itemize}

@@

\cref{fi:performance_study_load} shows the summary of the load testing plan with changing the request size and performing load testing with increasing the \acrshort{rps} over time. As shown in the figure, for a given request size, the \acrshort{rps} increased in steps over the elapsed time. In each step, the \acrshort{rps} holds for a moment to give some time for the system to get stabilized. Thus at each step, we can measure the performance of the system and study the stability of the system. After the test plan comes to the peak \acrshort{rps}, the system holds for more time than the previous steps. By holding the peak load for some time, we are planning to show the system is capable of scaling more than the proposed peak load in the testing plan. During the WDIAS load testing, we are planning to have a peak of 300 requests per second at peak time. After the peak \acrshort{rps} is over, the test case goes through a cool-down period to measure the ability to shrink down the system resources when there is not any load on the system.

@@

During the performance analysis, we used the following metrics to measure the performance of the system.
\begin{itemize}
    \item \emph{Throughput} -- Number of requests that can be processed per unit time. 
    Normally for \acrshort{wdias} performance observations, it uses \acrfull{rps} to measure the throughput.
    \item \emph{Latency} -- Taken time to respond to the request sent to the system. 
    Normally we measure with milliseconds (ms) in the \acrshort{wdias}. During the performance analysis, we measure the minimum response time, maximum response time, the average response time, and standard deviation. The standard deviation measures the mean distance of the values to their average value. It gives an idea of the dispersion or variability of the measures to their mean value. So, if the deviation value is low compared to the mean value, it will indicate that the measures are not dispersed (or mostly close to the mean value), and the mean value is significant. Other than that, we measure the response time into 90\% percentile while giving the best response time up to 90\% percent of the responses.
    \item \emph{Resource utilization} -- Resource Utilization is calculated based on the CPU and Memory usage via the K8s metrics server. It calculates the resource utilization over a 60 seconds time window. Network usage is getting based on the actual physical machines.
    \item \emph{Auto-scaling} -- We performed the load testing with two system configurations. First, we ran the microservices with a fixed number of pods inside the K8s cluster. Then k8s are only allowed to vertically scale each microservice until a container hits the CPU and memory limits defined per container or overwhelm the physical node resources. Secondly, we ran the load testing with auto-scaling enabled only for microservices, which are using higher resources with the above system configurations. Auto-scaling spawns new copies of the same microservice when the microservice gets more load than the configured percentage, and shrink the running instances when there is a lesser load.
\end{itemize}

>>>>

\section{Test Plan}
\label{se:test_plan}

In \cref{se:high_level_design} we discussed three modules of weather data integration and assimilation systems such as integration, assimilation, and dissemination. We can map each of those modules to WDIAS modules, which provide the same functionalities as below.
\begin{enumerate}
    \item \emph{Import} modules - Integration modules
    \item \emph{Export} modules - Dissemination modules
    \item \emph{Extension} modules - Assimilation modules
\end{enumerate}

@@

A timeseries is a series of data points in time order. The system may receive the data points of a timeseries in multiple requests. The request size can vary based on the use case from one data point to multiple data points. For example;
\begin{itemize}
    \item A weather station may send the current water level in a regular 1 minute intervals. In such a case, it always comes as a request with one data point with a higher frequency.
    \item Due to resource issues such as batteries, a weather station can collect data points for a given period and send as a batch. An example scenario would be, a weather station can collect the precipitation at each 5 minute period and send all the data after collecting for one hour.
    \item The forecast precipitation data from the \acrshort{wrf} model extract for a few days in hourly intervals and store them.
\end{itemize}

The percentage of number of requests per each data type that comes to the system at a given time can be varied depending on the situation. The Majority of weather data requests come from weather stations and other observation tools. Then we need to validate and modify those timeseries data and store them as new timeseries that can directly feed into the NWMs. These weather stations can send the weather data at different frequencies with different parameters such as precipitation, temperature, humidity, and wind direction. Most of the timeseries mentioned above are scalar timeseries. When compared to scalar data, parameters such as wind direction are vector timeseries with a lesser percentage of requests. On the other hand, most of the grid type data was generated as results of NWMs. But those grid data produce in less frequency due to the cost of the computer resource to produce results. After considering those factors we came up with 70\% scalar, 20\% vector, and 10\% grid percentage values as the expected requests rate ratios.
As an upper limit of such scenarios while designing the performance testing, it uses the following;
\begin{itemize}
    \item Hourly (24 data points)
    \item Every 30 minutes (48 data points)
    \item Every 15 minutes (96 data points) - One request is equivalent to a set of data points sent by a weather station in 15 minutes interval for a single day
\end{itemize}

@@

\cref{fi:performance_study_load} shows the summary of the load testing plan with changing the request size and performing load testing with increasing the \acrshort{rps} over time. As shown in the figure, for a given request size, the \acrshort{rps} increased in steps over the elapsed time. In each step, the \acrshort{rps} holds for a moment to give some time for the system to get stabilized. At each step, we can measure the performance of the system and study the stability of the system. After the test plan comes to the peak \acrshort{rps}, the system holds for more time than the previous steps. By holding the peak load for some time, we are planning to show the system is capable of scaling more than the proposed peak load in the testing plan. During the WDIAS load testing, we are planning to have a peak of 300 requests per second at peak time. After the peak \acrshort{rps} is over, the test case goes through a cool-down period to measure the ability to shrink down the system resources when there is not any load on the system.

@@

During the performance analysis, we used the following metrics to measure the performance of the system.
\begin{itemize}
    \item \emph{Throughput} -- Number of requests that can be processed per unit time. 
    Normally for \acrshort{wdias} performance observations, it uses \acrfull{rps} to measure the throughput.
    \item \emph{Latency} -- Taken time to respond to the request sent to the system. 
    Normally we measure with milliseconds (ms) in the \acrshort{wdias}. During the performance analysis, we measure the minimum response time, maximum response time, the average response time, and standard deviation. The standard deviation measures the mean distance of the values to their average value. It gives an idea of the dispersion or variability of the measures to their mean value. So, if the deviation value is low compared to the mean value, it will indicate that the measures are not dispersed (or mostly close to the mean value), and the mean value is significant. Other than that, we measure the response time into 90\% percentile while giving the best response time up to 90\% percentile of the responses.
    \item \emph{Resource utilization} -- Resource Utilization is calculated based on the CPU and Memory usage via the K8s metrics server. It calculates the resource utilization over a 60 seconds time window. Network usage is getting based on the actual physical machines.
    \item \emph{Auto-scaling} -- We performed the load testing with two system configurations. First, we ran the microservices with a fixed number of pods inside the K8s cluster. Then k8s are only allowed to vertically scale each microservice until a container hits the CPU and memory limits defined per container or overwhelm the physical node resources. Secondly, we ran the load testing with auto-scaling enabled only for microservices, which are using higher resources with the above system configurations. Auto-scaling spawns new copies of the same microservice when the microservice gets more load than the configured percentage, and shrink the running instances when there is a lesser load.
\end{itemize}

### 2020-05-08 results/sec_2-workload.tex
v5.1.1

We used the Apache JMeter for performing the load test since it is open source and supports distributed load testing. As mentioned in the JMeter Best Practices, "If your test needs large amounts of data - particularly if it needs to be \emph{randomized}, create the test data in a file that can be read with CSV Data set. This avoids wasting resources at run-time". While running the load test, JMeter iterates over the timeseries metadata CSV file and assigns a new thread to work on the test scenario with given timeseries metadata. Instead of providing the randomness at run time, we stored the timeseries metadata in random order that satisfies the percentage of 70\% scalar, 20\% vector, and 10\% grid, as mentioned in \cref{se:test_plan}.
Before running the test cases, we generated the timeseries metadata in random order and wrote them into a CSV file during the setup phase. The metadata CSV file arranges in a way that, within every ten lines, seven lines are Scalar timeseries, two lines are Vector timeseries, and 1 line is a Grid timeseries.

\acrshort{wdias} is capable of storing any numeric value up to 3 decimal points as per the default configuration. Even the parameter type is either precipitation or temperature does not make any difference, because in storage perspective, it's a number. Thus, during the performance testing, we only use real precipitation data from \acrshort{curw}. So, we used data from five weather stations for a month while preparing the test data for scalar timeseries. We used a script to clean up and prepare the precipitation data, and it is available as setup precipitation in \acrshort{wdias} codebase.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{results/work_load/experimental_setup_v3.jpg}
    \caption{Experimental setup with JMeter}
    \label{fi:experimental_setup}
\end{figure}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.75\textwidth]{results/work_load/eks_node_setup.jpg}
    \caption{\acrfull{eks} node setup}
    \label{fi:eks_node_setup}
\end{figure}

>>>>

We used the Apache JMeter for performing the load test since it is open source. And supports distributed load testing. As mentioned in the JMeter Best Practices, "If your test needs large amounts of data - particularly if it needs to be \emph{randomized}, create the test data in a file that can be read with CSV Data set. This avoids wasting resources at run-time". While running the load test, JMeter iterates over the timeseries metadata CSV file and assigns a new thread to work on the test scenario with given timeseries metadata. Instead of providing the randomness at run time, we stored the timeseries metadata in random order that satisfies the percentage of 70\% scalar, 20\% vector, and 10\% grid, as mentioned in \cref{se:test_plan}.
Before running the test cases, we generated the timeseries metadata in random order and wrote them into a CSV file during the setup phase. The metadata CSV file arranges in a way that, within every ten lines, seven lines are Scalar timeseries, two lines are Vector timeseries, and 1 line is a Grid timeseries.

\acrshort{wdias} is capable of storing any numeric value up to 3 decimal points as per the default configuration. Even the parameter type is either precipitation, temperature does not make any difference, because in storage perspective, we are storing numeric value. Thus, during the performance testing, we only use real precipitation data from \acrshort{curw}. So, we used data from five weather stations for a month while preparing the test data for scalar timeseries. We used a script to clean up and prepare the precipitation data, and it is available as setup precipitation in \acrshort{wdias} codebase.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.6\textwidth]{results/work_load/experimental_setup_v3.pdf}
    \caption{Experimental setup with JMeter}
    \label{fi:experimental_setup}
\end{figure}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.75\textwidth]{results/work_load/eks_node_setup.pdf}
    \caption{\acrfull{eks} node setup}
    \label{fi:eks_node_setup}
\end{figure}
